\documentclass{article}

\title{%
	A machine learning system for short-term weather prediction \\ \, \\
	\large Draft}

\begin{document}
	
	\maketitle 
	\section{Introduction}
	\subsection{Abstract}
	Short introduction to a problem and proposed solution.
	\subsection{Motivation}
	Emphasizing the exponential growth of data worldwide and comparison of NWP and MLP approaches with the suggestion that machine learning methods are especially promising for future research and development.
	\subsection{Goals}
	Obtaining sufficient results that allows to confirm thesis that Graph Neural Networks and generally machine learning approaches for weather prediction task is promising and worth developing.
	\subsection{Scope}
	\subsubsection{Model}
	Propose lighter version of GraphCast architecture and compare it with other solutions such as Convolutional Neural Networks, methods based on fluid dynamics or other related to Numerical Weather Prediction approach etc.
	\subsubsection{App}
	Create a robust and reliable application with API that communicate with our model.
	\subsection{Methodology}
	Definition of Agile and description of launched habits. 
	
	\section{Theoretical introduction}
	\subsection{Problem introduction}
	Theoretical introduction to a problem and mathematical concepts. 
	\subsection{Neural network concept}
	Explaining architecture of dense neural networks and other types such as convolutional, recurrent, graph (depends which one eventually will be used). 
	\subsection{Optimization algorithms}
	Short introduction to gradient descent and backpropagation algortihms. 
	\subsection{GraphCast}
	Deep dive into GraphCast architecture.
	\subsection{Model}
	Proposed model architecture and comparision with GraphCast. 
	\section{Report and analysis}
	\subsection{Data source and EDA}
	Description of dataset, preprocessing, visualisation and analysis.
	\subsection{Experiments and results}
	Results presentation, comparision with other models/benchmark scores. 
	\subsection{Practical details}
	Explain training, evaluating and forecasting processes;
	regularization techniques, metrics, post processing etc.
	\section{Application}
	\subsection{Functionality}
	\subsection{API}
	\subsection{Data Storage and DevOps solutions}
	
	\section{Conclusions}
	Limatation of proposed solution, plans for future development and conclusions.
	
	\section{Literature}
	Bibliography, referenced research papers.
	
\end{document}

